\actTitle{1.1 - Rectangular Coordinate System}



\noindent \textbf{Topics:}  $x,y$-plane and coordinate system, quadrants, plotting points, distance formula, graphing an equation, $x$ and $y$ intercepts\\

\noindent \textbf{Student Learning Outcomes:}
\begin{enumerate}
\item Students will be able to plot points in an $x,y$-plane using a coordinate system.
\item Students will be able to determine the distance between two points on an $x,y$-plane using the distances formula.
\item Students will be able to graph an equation on an $x,y$-plane and determine the $x$ and $y$ intercepts of the equation.
\end{enumerate}

\hrule
\bigskip

\noindent \textbf{Rectangular Coordinate System}\\
\begin{center}
\begin{tikzpicture}
\begin{axis}[
    xmin=-5, xmax=5,
    ymin=-5, ymax=5,
    axis lines=center,
    axis on top=true,
    domain=0:1,
    ]

   
\end{axis}
\end{tikzpicture}
\end{center}


\begin{enumerate}
\item Plot and label the given points on the coordinate axes below.\\

$$A(2,-3) \quad \quad B(-2,0) \quad \quad C(-1,2)$$
\begin{center}
\begin{tikzpicture}
\begin{axis}[
    xmin=-5, xmax=5,
    ymin=-5, ymax=5,
    axis lines=center,
    axis on top=true,
    domain=0:1,
    ]

   
\end{axis}
\end{tikzpicture}
\end{center}

\newpage
\noindent \textbf{The Distance Formula}\\
\item Label the points $(1,5)$ and $(4,9)$ on the graph below.  Then draw a line representing the distance between the points and label that line $d$.

\begin{center}

\begin{tikzpicture}
\begin{axis}[
    xmin=-2, xmax=6,
    ymin=-2, ymax=10,
    axis lines=center,
    axis on top=true,
    domain=1:4,
    ]

\end{axis}

\end{tikzpicture}
\end{center}

\begin{enumerate}
\item Determine the horizontal distance between the points $(1,5)$ and $(4,9)$:\\
(This is the distance between the $x$-values.)\\[.3in]
\item Determine the vertical distance between the points $(1,5)$ and $(4,9)$:\\
(This is the distance between the $y$-values.)\\[.3in]
\item How can we use these two values to determine the straight-line distance $d$ between the points $(1,5)$ and $(4,9)$?\\[.3in]
\item Calculate the distance between the points $(1,5)$ and $(4,9)$.

\end{enumerate}
\vfill

\noindent \textbf{The Distance Formula:}  The distance between points $(x_1,y_1$ and $(x_2,y_2)$ is given by $$d=\sqrt{(x_2-x_1)^2+(y_2-y_1)^2}$$

\newpage

\noindent \textbf{Graph Equations by Plotting Points}\\
\item Graph the equation $y=x+2$.\\
\begin{center}
\begin{tikzpicture}
\begin{axis}[
    xmin=-5, xmax=5,
    ymin=-5, ymax=5,
    axis lines=center,
    axis on top=true,
    domain=0:1,
    ]

   
\end{axis}
\end{tikzpicture}
\end{center}

\item Graph the equation $y-|x|=-1$.
\begin{center}
\begin{tikzpicture}
\begin{axis}[
    xmin=-5, xmax=5,
    ymin=-5, ymax=5,
    axis lines=center,
    axis on top=true,
    domain=0:1,
    ]

   
\end{axis}
\end{tikzpicture}
\end{center}


\noindent \textbf{Finding $x$ and $y$ Intercepts}\\
An $x$-intercept is a point $(x,y)$ where a graph crosses the $x$-axis and a $y$-intercept is a point $(x,y)$ where a graph crosses the $y$-axis.\\

\item Determine the $x$-intercept(s) and $y$-intercepts of the previous example, $y-|x|=-1$ and label them on the graph.
\begin{enumerate}
\item $x$-intercept(s):\\
$y$-intercept(s):\\
\item What do you notice about the intercepts?\\
\item How are they different?
\end{enumerate}
\newpage

\item Determine the $x$-intercepts of $y-|x|=-1$ algebraically.
\vfill
\item Determine the $y$-intercept of $y-|x|=-1$ algebraically.
\vfill
\end{enumerate}

\noindent \textbf{Student Learning Outcomes Check}

\begin{enumerate}
\item Can you plot points in an $x,y$-plane using a coordinate system?\\
\item Are you able to determine the distance between two points on an $x,y$-plane using the distances formula?\\
\item Can you graph an equation on an $x,y$-plane and determine the $x$ and $y$ intercepts of the equation?
\end{enumerate}

\noindent \textbf{If any of your answers were no, please ask about these topics in class.}


