\documentclass[11pt]{article}
\usepackage{amsmath, amssymb, array}
\pagestyle{plain}

\textwidth=6.5in
\hoffset-1in
\textheight=9in
\voffset-1in

\usepackage{enumitem}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage{lipsum}
\usepackage{stfloats}
\usepackage{multicol}
\setlength{\columnsep}{1cm}




\begin{document}


\noindent MATH 1113   \hfill 2.1 Part 2 - Quadratic Functions and Applications\\



\noindent \textbf{Topics:}  quadratic functions, applications of quadratic functions\\

\noindent \textbf{Student Learning Outcomes:}
\begin{enumerate}
\item Students will be able to determine a quadratic function given a description.
\item Students will be able to apply a quadratic function to geometry.
\end{enumerate}

\hrule 
\vspace{5mm}
\section{Solve Applications Involving Quadratic Functions}


\begin{enumerate}
\item A leaping cat follows a parabolic path. The cat jumps a maximum height of 5 feet and covers 6 feet of horizontal ground distance. Find an equation of the form $y = a(x-h)^2+k$ which represents the path of the cat. \vfill

\item An object is projected vertically upward from the top of a building with an initial velocity of 112 ft/sec. Its distance $s(t)$ in feet above the ground after $t$ seconds is given by the equation $s(t)=-16t^2 + 112t + 110.$
\begin{enumerate}
\item Find its maximum distance above the ground.
\item Find the height of the building.
\end{enumerate}

\vfill



\newpage


\section{Applying a Quadratic Function to Geometry}


\item A parking area is to be constructed adjacent to a road.  The developer has purchased 340 ft of fencing.  Determine dimensions for the parking lot that would maximize the area.  Then find the maximum area.


\vfill



\end{enumerate}

\noindent \textbf{Student Learning Outcomes Check}

\begin{enumerate}
\item Can you determine a quadratic function given a description?
\item Are you able to apply a quadratic function to geometry?

\end{enumerate}

\noindent \textbf{If any of your answers were no, please ask about these topics in class.}













\end{document}