\documentclass[11pt]{article}
\usepackage{amsmath, amssymb, array}
\pagestyle{plain}

\textwidth=6.5in
\hoffset-1in
\textheight=9in
\voffset-1in

\usepackage{enumitem}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage{lipsum}
\usepackage{stfloats}
\usepackage{multicol}
\usepackage{minipage-marginpar}
\usepackage{tikz}
\setlength{\columnsep}{1cm}

\newcommand{\abs}[1]{\left| #1 \right|}


\begin{document}


\noindent MATH 1113   \quad\quad\quad\quad\quad Worksheet \quad\quad\quad\quad\quad\   Name \underline{\phantom{alphabetsoupismyveryveryfavorite}}\\ 
\noindent Section 3.2-3.4 \\




\noindent \textbf{Instructions:}  Work together in groups of  3 or 4 to complete the following problems.



\begin{enumerate}

\item Which functions are exponential functions?
$$f(x)=4.2^x \quad \quad g(x)=x^{4.2} \quad \quad h(x)=4.2x \quad \quad k(x)=(\sqrt{4.2})^x \quad \quad m(x)=(-4.2)^x$$

\item Consider $y=3^x$.  Determine the domain, range, and asymptote for each of the following functions.

\begin{enumerate}
\item $f(x)=3^x$\vfill
\item $g(x)=3^x+2$\vfill
\item $b(x)=3^{x+2}-1$\vfill


\item  $\displaystyle h(x)=\Big(\frac{1}{3}\Big)^x$
\vfill


\end{enumerate}

\item Alice needs to borrow \$15,000 to buy a car.  She can borrow the money at 6.4\% compounded monthly for 5 years or she can borrow the money at 6.7\% interest compounded continuously for 5 years.  Which option is most cost effective for Alice?

\vfill
\vfill

\newpage

\item Write the domain in interval notation.
\begin{enumerate}
\item $f(x)=\log(8-x)$ 
\vfill
\item $g(x)=\log_2 (x^2-16)$ 
\vfill
\item $h(x)=\ln(x^2+14)$
\vfill
\item $\displaystyle m(x)=3+\log_4\Bigg(\frac{1}{\sqrt{11-x}}\Bigg)$
\vfill
\end{enumerate}



\item Simplify the expression without using a calculator.
\begin{enumerate}

\item $\log_3(9)$\\

\item $\displaystyle \log_2(\frac{1}{16})$\\

\item $\log_{1/7}(49)$\\



\end{enumerate}




\newpage


\item Simplify the expression without using a calculator.
\begin{enumerate}

\item $\displaystyle \log_4(4^{11})$\\

\item $\displaystyle 5^{\log_5(x+y)}$\\

\item $\log_{\pi}(1)$\\



\end{enumerate}






\item Write the logarithm as a sum or difference of logarithms and simplify as much as possible.  (Expand the logarithmic expression.)
\begin{enumerate}
\item $ \displaystyle \log_7\Big(\frac{1}{7}mn^2\Big)$ 
\vfill
\item $\displaystyle \log_5\Big(\frac{p^5}{mn}\Big)$ 
\vfill

\item $\displaystyle \log\Bigg(\frac{10}{\sqrt{a^2+b^2}}\Bigg)$ 
\vfill

\item $\displaystyle \ln\Bigg(\sqrt[5]{\frac{e^2}{c^2+5}}\Bigg)$ 
\vfill


\end{enumerate}

\newpage

\item Write the logarithmic expression as a single logarithm with coefficient 1, and simplify as much as possible.  (Condense the logarithmic expression.)


\begin{enumerate}
\item $\ln(y)+\ln(4)$\vfill
\item $\log_3(693)-\log_3(33)-\log_3(7)$\vfill
\item $3[\ln(x)-\ln(x+3)-\ln(x-3)]$\vfill
\item $\displaystyle 15\log(c)-\frac{1}{4}\log(d)-\frac{3}{4}\log(k)$\vfill
\end{enumerate}




\newpage

\item Use the change-of-base formula to write the following as a single logarithm. $$(\log_3(11))(\log_{11}(4))$$ 


\vfill
\item  Show that $$-\ln(x-\sqrt{x^2-1})=\ln(x+\sqrt{x^2-1})$$

  
\vfill

\end{enumerate}


\end{document}