%=========================================================================
% Start of
%=========================================================================
\preClass{Introduction to Logarithms}

\begin{problem}
\item A computer virus is constructed that will infect two new
  computer systems each day and then wipe the disk drive of its
  current computer clean.  The virus is installed on one computer on
  1 January.
  \begin{subproblem}
  \item How many new computers will it infect on 2 January?
    \vfill
  \item How many new computers will it infect on 3 January?
    \vfill
  \item How many new computers will it infect on 4 January?
    \vfill
  \end{subproblem}
\item A computer virus is constructed that will infect two new
  computer systems each day and then wipe the disk drive of its
  current computer clean.  The virus is installed on a computer but it
  is not clear when it was first installed.
  \begin{subproblem}
  \item It is estimated that 32 computers were infected on a given
    day. How many days beforehand was it installed on the first computer?
    \vfill
  \item It is estimated that 64 computers were infected on a given
    day. How many days beforehand was it installed on the first computer?
    \vfill
  \item It is estimated that 512 computers were infected on a given
    day. How many days beforehand was it installed on the first computer?
    \vfill
  \end{subproblem}
  \clearpage
\item Make a sketch of the number line. Mark zero on your number
  line. Indicate the locations of 1, 1.5, -0.5, and 9.5. Try to make
  the points consistent with the relative distances between the
  points. Also, label the number line.
\end{problem}


\actTitle{Introduction to Logarithms}
\begin{problem}
\item The approximate distances in kilometers from the earth to
  various destinations is given in the table below. Use this
  information to answer each of the questions below.

  \begin{tabular}{ll}
    Destination & Distance \\ \hline
    The moon & $3.84 \times 10^5$ km \\
    Mars &  $5.46 \times 10^7$ km \\
    Saturn & $1.28 \times 10^9$ km \\
    Proxima Centauri (Closest star) & $4.01 \times 10^{13}$ km \\
    Center of the Milky Way & $2.83 \times 10^{17}$ km
  \end{tabular}

  \begin{subproblem}
  \item Make a sketch of a number line with the distances for each
    destination marked on the number line.  Try to keep the relevant
    distances between points consistent.

    \vfill
    
  \item For each distance in the table above what is the approximate
    power of ten for the distance? For example, if a distance is
    $1\times 10^4$ km it is a power of 4. Sketch a number line and
    indicate the powers of ten on the number line. Try to keep the
    relevant distances between points consistent.
      
    \vfill
  \end{subproblem}

\clearpage

The approximate average lengths of various items are given in the table
below. Use this information to answer each of the questions below.

  \begin{tabular}{ll}
    Item & Size \\ \hline
    Person & 1.75 m \\
    Finger & $9.2 \times 10^{-2}$ m\\
    DNA & $ 5.0 \times 10^{-2}$ m \\
    Hair (Width) & $ 9.0\times 10^{-5}$ m \\
    Bacteria & $ 0.8 \times 10^{-7}$ m\\
  \end{tabular}

  \begin{subproblem}
    \item Make a sketch of a number line with the lengths for each
      item  marked on the number line.
      \sideNote{Try to keep the relevant distances between points consistent.}
      \vfill
    \item For each item in the table above what is the approximate
      power of ten for the length? For example, if a length is
      $1\times 10^{-4}$ m it is a power of -4. Sketch a number line
      and indicate the powers of ten on the number line.
      \sideNote{Try to keep the relevant distances between points consistent.}
      \vfill
  \end{subproblem}

\clearpage

\item In one cycle a type of mayfly lays eggs and on average three
  females survive to lay more eggs. One of the mayfly are introduced
  to a stream for the first time and lays eggs.
  \begin{subproblem}
  \item Draw a tree diagram of the individuals from each cycle that
    survive starting with the original mother. Include three cycles in
    your diagram. The node for the first female mayfly should be at
    the top center of the space, and then the generations that follow
    should be below the previous generations in the diagram.  

    \vfill

  \item On the right side of the tree diagram indicate the total number
    of female mayflies that survive each cycle.
  \item On the left side of the tree diagram indicate the coresponding cycle with the
    first cycle labeled as cycle 0 (zero), and the second cycle is 1 (one).
  \item It is estimated that there are 2187 female mayfly in the river. How
    many cycles have there been?
    \vspace{3em}
  \item It is estimated that there are 19683 female mayfly in the river. How
    many cycles have there been?
    \vspace{3em}
  \item If the number of female mayfly, $M$, is given by $M=3^n$ where
    $n$ is the number of cycles, solve the equation for $n$. (What is
    your new equation called in relation to the original function?)
    \vspace{3em}
  \end{subproblem}

\clearpage

\item Mosquitoes lay between 50 to 200 eggs each cycle. Assume that
  for a given species in a particular area roughly 34 eggs hatch and
  survive to be female adults and lay eggs.
  \begin{subproblem}
  \item Assume that in the spring there is one surviving female
    mosquito starting with cycle 0. Make a \textbf{table} to indicate
    how many mosquitoes there will be from cycle 0 to cycle 4.

    \vfill
    
  \item Using your table, if it is estimated that there are 39304 female
    mosquitoes what cycle in the season is it?
    
    \vfill
    
  \item Using your table, if it is estimated that there are 1,336,336 female
    mosquitoes what cycle in the season is it?
    
    \vfill
    
  \item If the number of female mosquitoes, $f$, is given by $f=34^n$ where
    $n$ is the number of cycles, solve the equation for $n$. (What is
    this equation called in relation to the original function?)
    \vspace{3em}
  \end{subproblem}

\clearpage

\item Determine the value(s) of $x$ that satisfy the equation
  \begin{eqnarray*}
    \log_5(x) + \log_5(x+1) & = & 2.
  \end{eqnarray*}

  \begin{subproblem}
  \item Raise both sides of the equation to the power of 5. 
    \vfill
  \item Take advantage of the property that $5^{a+b}=5^a\cdot 5^b$ to
    rewrite the left hand side as the product of two values.
    \vfill
  \item Use the property of the inverse to rewrite each of the terms
    in a simpler form.
    \vfill
  \item Simplify the right hand side and solve for $x$.
    \vfill
  \end{subproblem}

\end{problem}

\postClass

\begin{problem}
\item Briefly state two ideas from today's class.
  \begin{itemize}
  \item
  \item
  \end{itemize}
\item Determine the value of $x$ in each expression below. Any
  approximations should be to at least two decimal places.
  \begin{subproblem}
    \item $\log_5(2x+1) = 9$
    \item $\log_{7}(x-1) + \log_{7}(x+1) = 2$
    \item $3^{\log_3(x+1)} = 4$
    \item $\log_8\left(8^{5-2x}\right) = 9$
    \item $\log_{10}(2x+1) = \log_3(4.1)$
  \end{subproblem}
\item Use the natural logarithm to determine the value of $x$ in each
  expression below. Any approximations should be to at least two
  decimal places.
  \begin{subproblem}
  \item $3.5^{x+1}=2^{x-1}$
  \item $10=e^{-2x}$
  \item $1=2^{3x}\cdot 4^{8x-1}$
  \item $5=e^{3x-1}$
  \item $e^{9x} = 2^{9x}$
  \end{subproblem}
\end{problem}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../labManual"
%%% End:
