
\actTitle{Worksheet 3.1}

\noindent \textbf{Instructions:}  Work together in groups of  3 or 4 to complete the following problems.


Student goals:
\begin{itemize}
\item Determine if a given function is one-to-one. The function can
  be given in graphical, tabular, or algebraic forms.
\item Explicitly show that a function is one-to-one or demonstrate why
  a function is not one-to-one.
\item Determine the inverse of a given function. The function can be
  given in tabular or algebraic forms.
\item Determine the domain and range of the inverse of a function. The
  function can be given in graphical, tabular, or algebraic forms.
\item Limit the domain of a function that is not one-to-one so that an
  inverse can be defined on the resulting restricted domain.
\end{itemize}


\begin{enumerate}
\item Choose 3 of the following functions for your group to work with.  You must choose $x^2$ and two other functions.
$$f(x)=x \quad \quad g(x)=x^2 \quad \quad h(x)=x^3 \quad \quad j(x)=\frac{1}{x} \quad \quad m(x)=\sqrt{x}  \quad p(x)=\sqrt[3]{x}$$

\item Use transformations like shifting, stretching, compressing, and reflecting to transform each of your 3 functions.  Then write each transformed function.

\item Rotate to a new set of functions.  Then \emph{algebraically} determine if your new functions are one-to-one.  Label each function as \textbf{one-to-one} or \textbf{not one-to-one}.

\item Rotate to a new set of functions.  For your new functions that are one-to-one, find their inverse functions.  

\item Rotate to a new set of functions.  Verify that the original function and it's inverse are, in fact, inverses of each other by \textbf{function composition}.

\item Work on your own and choose two other functions from \#1 to transform and work through this process of determining if it is one-to-one, finding it's inverse function, and verifying the inverse by function composition.




\end{enumerate}




