%=========================================================================
% Start of
%=========================================================================
\preClass{The Natural Logarithm}

\begin{problem}
\item Determine an approximation for the numerical value of each
  number below. Also determine the natural logarithm of each of the
  following values. Express the numbers to two decimal places.
  \begin{subproblem}
  \item $e$
    \vfill
  \item $e^2$
    \vfill
  \item $e^3\cdot e^{-4}$
    \vfill
  \item $e^{-2}\cdot e^{-3}$
    \vfill
  \end{subproblem}
\end{problem}


\actTitle{The Natural Logarithm}
\begin{problem}
\item Determine the natural logarithm of each value below, and use the
  properties of logarithms to express the value as a sum or
  differences of logarithms.
  \begin{subproblem}
  \item $a\cdot b \cdot c$
    \vfill
  \item $a\cdot b^2 \cdot c \cdot d^3$
    \vfill
  \item $\frac{(x-1)\cdot (x+3)}{(x-2)}$
    \vfill
  \item $\frac{(x-4)\cdot (x+2)^3 \cdot (x+4)}{(x-9)^2}$
    \vfill
  \end{subproblem}

\clearpage

\item A computer virus is constructed that will infect two new
  computer systems each day and then wipe the disk drive of its
  current computer clean.  The virus is installed on one computer on
  day 0 (zero).
  \begin{subproblem}
    \item Determine the formula that gives the number of computers
      infected on the number of days since the first computer is
      infected.
      \begin{eqnarray*}
        \mathrm{Number(t)} & = &
      \end{eqnarray*}
    \item Make a rough sketch of the number of new computers infected
      as a function of time.
      \sideNote{Label your axes and annotate your plot.}
      \vfill
    \item Solve the equation above for $t$. That is determine a formula for
      $t$ given the number of newly infected computers.
      \vfill
    \item It is determined that 524288 new computers were infected on
      a given day. How long ago was the first virus installed?
      \vspace{3em}
    \item It is determined that 67108864 new computers were infected
      on a given day. How long ago was the first virus installed?
      \vspace{3em}
  \end{subproblem}

\clearpage

\item Radon 222 has a half life of 3.8 days. It is estimated that
  there is 0.2g of radon 222 in a basement. Assume that no more
  radon enters the basement after a treatment is applied, and the area
  is not ventilated.
  \begin{subproblem}
    \item Determine a formula for the amount of radon in the basement
      at a given time, in days, from when the treatment is applied.
      \begin{eqnarray*}
        \mathrm{Amount(t)} & = &
      \end{eqnarray*}
    \item Make a rough sketch of the amount of Radon 222 in the
      basement as a function of time.
      \sideNote{Label your axes and annotate your plot.}
      \vfill
    \item Solve the original equation for $t$. That is determine a
      formula for $t$ given the amount of radon in the basement.
      \vfill
    \item How long will it take before the amount of radon is down to
      one tenth the original amount?
      \vfill
  \end{subproblem}

\clearpage

\item \label{question:massOfGrowth} A patient has a growth, and a
  treatment is applied that is estimated to reduce the mass of the
  growth by ten percent each week. At the start of the treatment the
  mass of the growth is estimated to be two grams.
  \begin{subproblem}
  \item Determine a formula for the mass of the growth at a given
    time, in days, from when the treatment is applied.
      \begin{eqnarray*}
        \mathrm{Size(t)} & = &
      \end{eqnarray*}
      \vfill
      \vfill
    \item Make a rough sketch of the mass of the growth as a function
      of time.
      \sideNote{Label your axes and annotate your plot.}
      
      \vfill
      \clearpage
      
    \item Solve the equation for $t$. That is determine a formula for
      $t$ given the size of the growth.
      
      \vfill
      
    \item How long will it take before the growth is reduced to one
      tenth of its original mass?

      \vfill
      
  \end{subproblem}


\end{problem}


\postClass

\begin{problem}
\item Briefly state two ideas from today's class.
  \begin{itemize}
  \item
  \item
  \end{itemize}
\item Determine the value of $x$ in each expression below:
  \begin{subproblem}
  \item $e^{2x} - 2 e^{x} - 3 = 0$.
  \item $e^{2x} - 4 e^{x} + 4 = 0$.
  \item $e^{2x} -   e^{x} + 6 = 0$.
  \item $e^{2x} -   e^{x} - 12 = 0$.
  \item $e^{2x} - e^{x} + 4 = 0$.
  \end{subproblem}
\item Complete question \ref{question:massOfGrowth}.
\item A bank offers 1.5\% annual interest compounded weekly (assume 52
  weeks in a year). How long will it take for the balance to double?
\item A bank offers 1.5\% annual interest compounded monthly. How long
  will it take for the balance to double?
\item A compound is created that decays over time. It takes four years
  until half of the compound decays in a sample. How long will it take
  for 80\% of the compound to decay?
\item A species of plant produces two hundred seeds each year, but on average only 5\%
  of the seeds germinate and grow in to plants that produce seeds. A survey
  is done in a large area and it is estimated that the area contains fifty
  plants. How many plants are expected to be present after 3 years? Determine
  a formula to estimate the number of plants at any year, $t$, in the future.
\end{problem}



\propertiesTitle{Properties of logarithms}

Logarithmic functions satisfy the algebraic properties given below. In
each example it is assumed that $a$, $b$, and $r$ are constants. Both
$a$ and $b$ must be positive numbers.
\begin{eqnarray}
  \ln(a\cdot b) & = & \ln(a) + \ln(b) \\ [10pt]
  \ln\left(\frac{a}{b}\right) & = & \ln(a) - \ln(b) \\  [10pt]
  \ln\left(a^r\right) & = & r \ln(a) \\ [30pt]
  \log_{10}(a\cdot b) & = & \log_{10}(a) + \log_{10}(b) \\ [10pt]
  \log_{10}\left(\frac{a}{b}\right) & = & \log_{10}(a) - \log_{10}(b) \\  [10pt]
  \log_{10}\left(a^r\right) & = & r \log_{10}(a)
\end{eqnarray}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../labManual"
%%% End:
