%=========================================================================
% Start of section on modeling with exponential and log equations.
%=========================================================================
\preClass{More Exponential and Logarithmic Equations}

\begin{problem}
\item Solve for the variable $x$ in each expression below.
  \begin{subproblem}
  \item $e^{4x-1}  =  8$
    \vfill
  \item $\ln(2x+1)  =  -4$
    \vfill
  \item $3^{2x+1}  =  e^{4x}$
    \vfill
  \item $8^{3x+2}  =  6e^{4x}$
    \vfill
  \end{subproblem}
\end{problem}


\actTitle{Exponential and Logarithmic Models}
\begin{problem}
\item The number of animals in a population of a small mammals follows
  a logistic function,
  \begin{eqnarray*}
    P(t) & = & \frac{10,000}{1+e^{-\frac{1}{2}t}},
  \end{eqnarray*}
  where $t$ is the number of years since the initial observation.
  \begin{subproblem}
  \item What is the initial population?
    \vspace{4em}
  \item How many animals will the population approach after a very
    long time?
    \vspace{4em}
  \item How long will it take for the number of animals to reach 75\%
    of its long term value?
    \vfill
  \item How long will it take for the number of animals to reach 80\%
    of its long term value?
    \vfill
  \end{subproblem}

\clearpage

\item A spill occurs and a chemical is introduced into a lake. The
  amount of chemical decays over time. 
  \begin{subproblem}
  \item In this case an exponential function should be used. Why?
    \vspace{2em}

  \item Should the rate, $r$, used in the exponential function be
    positive or negative? (Briefly explain how you arrive at your
    conclusion.)  
    \vspace{2em}

  \item Write the general form for the equation.
    \vspace{2em}

  \item At some time after the spill occurs it is estimated that there
    are 4,000 kg of the chemical in the lake. A month later it is
    estimated that there is 3,500 kg in the lake. What is the half
    life of the chemical?
    \begin{subsubproblem}
    \item Write out the equations that result from the given information.
      \vspace{2em}

    \item Use the two equations to solve for the value of the decay rate.
      \vfill
      \vfill

    \item Determine the half life of the chemical.
      \vfill
    \end{subsubproblem}

    
  \end{subproblem}

\clearpage

\item A population of bacteria doubles every four hours. How long does
  it take for the population to triple?
  \begin{subproblem}
  \item Is this decay or growth? Make a (very) rough sketch of the
    graph of the function you expect to get.
    \vspace{3em}
  \item Write out the information that is given to you.
    \vspace{3em}
  \item Write out a general form of a function that will model this
    situation.
    \vspace{1em}
  \item Use the given information and the function above to construct
    a system of equations.
    \vspace{2em}
  \item Solve the system of equations for the constants necessary to
    solve the problem.
    \vfill
  \item How long will it take for the population to triple?
    \vfill
  \end{subproblem}

\clearpage

\item Carbon 14 has a half life of 5,730 years. How long will it take
  for a sample to decay to two thirds of its original value?
  \sideNote{Use the same steps used in the previous question.}
  \vfill
\clearpage



\end{problem}

\postClass

\begin{problem}
\item Briefly state two ideas from today's class.
  \begin{itemize}
  \item
  \item
  \end{itemize}
\item The number of animals in a population is approximated using a
  logistic function,
  \begin{eqnarray*}
    P(t) & = & \frac{80,000}{1+10e^{-\frac{1}{10}t}}.
  \end{eqnarray*}
  \begin{subproblem}
  \item Make a sketch of the function for $t$ going from 0 to 50.
  \item How long will it take for the population to double with
    respect to its original population?
  \item How many animals does the population approach in the long run?
    (i.e. what happens to the population after a very long time?)
  \end{subproblem}
\end{problem}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../labManual"
%%% End:
