\preClass{Coordinate Systems}



\noindent Watch the Pre-Class videos for Section 2.1B and answer the following questions. Remember that in your written work you are graded on the correctness of your supporting work and not just your final answer. Always give an exact answer unless you are explicitly told to round; calculator approximations will not receive full credit. 


\begin{enumerate}

\item   You have a 500 foot roll of fencing and a large field.  You want to construct a rectangular playground area.  What are the dimensions of the largest such playground?  What is the largest area?
\begin{enumerate}
\item Draw a picture of the rectangular playground and label the side lengths using your own variables.
\vfill
\vfill
\item Using your variables, write an equation that represents the area of the playground.
\vfill
\item Your previous answer should have two variables.  Use 500 to represent the perimeter of the the playground and solve for one of your variables.  (Note: all four sides of the rectangle will be used.)
\vfill
\vfill
\item Rewrite your area equation in terms of only one variable and simplify.  The result should be a quadratic function.
\vfill


\newpage



\item To determine the maximum area possible for the playground, what part of the parabola do you need to locate?
\vfill

\item Find that part of the parabola (from your previous answer).
\vfill
\vfill
\item What is the largest area the playground could be?
\vfill

\item  What are the dimensions of that playground?
\vfill

\end{enumerate}







\end{enumerate}



