

\actTitle{Worksheet 1.5}


\noindent \textbf{Instructions:}  Work together in groups of  3 or 4 to complete the following problems.




\begin{enumerate}

\item A small business makes cookies and sells them at the farmer's
  market.  The fixed monthly cost for use of a Health
  Department-approved kitchen and rental space at the farmer's market
  is \$790.  The cost of labor, taxes, and ingredients for the cookies
  amounts to \$0.24 \textbf{per cookie}, and the cookies sell for
  \$6.00 \textbf{per dozen}.
\begin{enumerate}
\item Write a linear cost function representing the cost, $C(x)$
  dollars, to produce $x$ dozen cookies per month.  Then determine the
  monthly cost to produce 100 dozen cookies each month.
  \vfill
\item Write a linear revenue function representing the revenue, $R(x)$
  dollars, for selling $x$ dozen cookies.  Then determine the revenue
  for selling 100 dozen cookies.
  \vfill
\item Will the business make a profit or lose money if they sell 100 dozen cookies each month?
\vfill
\item Write a linear profit function representing the profit, $P(x)$,
  for producing and selling $x$ dozen cookies in a month. (Profit is
  revenue minus cost.)
  \vfill
\item Determine the least number of cookies (in dozens) that must be
  produced and sold to achieve a monthly profit.  Your answer should
  be an appropriate whole number.
  \vfill
\item If 150 dozen cookies are sold in a given month, how much money will the business make or lose?
\vfill
\end{enumerate}

\clearpage

\item The boiling point of water is 212$^\circ$ Fahrenheit which is
  equivalent to 100$^\circ$ Celsius. The freezing point of water is
  32$^\circ$ Fahrenheit, or 0$^\circ$ Celsius. Fahrenheit and Celsius
  are linearly related.
 
 \begin{enumerate}
 \item Determine a function that will tell you the temperature in degrees Celsius if
you already know the temperature in degrees Fahrenheit.
\vfill
\vfill
 \item If it is 83$^\circ$ Fahrenheit, determine the temperature in degrees Celsius.
 \end{enumerate}
 \vfill
 
 \item A car has a 15-gal tank for gasoline and gets 30 mpg on a highway while driving 60 mph. Suppose that the driver starts a trip with a full tank of gas and travels 450 mi on the highway at an average speed of 60 mph.
 
 \begin{enumerate}
 \item Write a linear model representing the amount of gas $G(t)$ left in the tank t hours into the trip.
 \vfill
 \vfill
 \item Evaluate $G(4.5)$ and interpret the meaning in the context of the problem.
 \vfill
 \end{enumerate}

\clearpage

\item A rectangle has a height, $h$, and a width, $w$. The perimeter
  of the rectangle is 40 cm. Determine the equation of $h$ as a
  function of $w$. Use your equation to determine how $h$ changes when
  $w$ is increased by 1 cm.

  \vfill

\item An animal pen is to be constructed in the shape of a rectangle
  with a height, $h$, and a width, $w$. The cost of the fencing that
  will be used to construct the pen is \$35.00 per meter. A total of
  \$20,000.00 is budgeted for fencing. Determine the equation of $h$
  as a function of $w$. Use your equation to determine how $h$ changes
  when $w$ is increased by 1 m.

  \vfill

 
 \clearpage
 
\item The table gives the number of calories and the amount of
  cholesterol for selected fast food hamburgers.
  \label{question:1.5:calories}
\begin{table}[h]
  \begin{center}
    \begin{tabular}{|l|l|}
      \hline
      \textbf{Hamburger Calories} & \textbf{Cholesterol (mg)} \\ \hline
      220                         & 35                        \\ \hline
      420                         & 50                        \\ \hline
      460                         & 50                        \\ \hline
      480                         & 60                        \\ \hline
      560                         & 70                        \\ \hline
      590                         & 105                       \\ \hline
      610                         & 65                        \\ \hline
      680                         & 80                        \\ \hline
      720                         & 90                        \\ \hline
    \end{tabular}
  \end{center}
\end{table}
\begin{enumerate}
\item Use the data points associated with 480 calories and 720
  calories to write a linear function that defines the amount of
  cholesteral, $chol(h)$, as a linear function of the number of
  hamburger calories $h$.
  \vfill
\item Interpret the meaning of the slope in the context of this problem.
\vfill
\item Use the model from part (a) to predict the amount of cholesterol for a hamburger with 650 calories.  
\vfill
\end{enumerate}


%\clearpage
%
%
%\item Given a square with sides of length $s$, diagonal of length $d$, perimeter $P$, and area $A$,
%\begin{enumerate}
%
%\item Write $P$ a function of $s$.
%\vfill
%\item Write $A$ as a function of $s$.
%\vfill
%\item Write $A$ as a function of $P$.
%\vfill
%\item Write $P$ as a function of $A$.
%\vfill
%\item Write $d$ as a function of $s$.
%\vfill
%\item Write $s$ as a function of $d$.
%\vfill
%\item Write $P$ as a function of $d$.
%\vfill
%\item Write $A$ as a function of $d$.
%\vfill
%
%\end{enumerate}
\end{enumerate}


\hwTitle{Section 1.5}

\begin{enumerate}
\item Finish question \ref{question:1.5:calories}.
\item It is determined that the rate of genetic change of the plants
  in a certain region is a linear function of the species richness of
  the region. When the richness is measured to be 110 species the
  genetic change is 300 mutations per year. It is estimated that an
  increase of 10 species results in an increase of 3 mutations per
  year. Determine the equation relating the genetic change and the
  species richness. (Explicitly define your variables and be
  consistent with your definitions.)
\item The Madison Rambler leaves the Chicago train station at
  8:00am. The train consumes 2,500 gallons of fuel every 500
  miles. When it leaves the station it has 4,000 gallons of fuel in
  its storage tanks. Determine the volume of fuel in the train's
  storage tanks as a function of the distance it has traveled from
  Chicago. Determine the domain of the function.
\item On a TV show, the contestants barter for individual bags of
  rice. Each day, they eat the same amount of rice. Six days after
  earning a bag of rice, one contestant has 23 oz remaining. Four days
  later, 17 oz rice remains in the person's bag. Determine the initial
  weight of the bag of rice and predict when it will run out.
\item An animal pen will be constructed along the bank of a river
  using two-hundred meters of fencing. The pen will be constructed in
  the shape of a rectangle. One side of the pen will border the river
  and will not require a fence. Determine the width of the pen as a
  function of its height.
\item An animal pen will be constructed along the bank of a river with
  a budget of \$5,000.00. The pen will be constructed in the shape of
  a rectangle, and fencing costs \$20.00 per meter. One side of the pen
  will border the river and will not require a fence. Determine the
  width of the pen as a function of its height.
\item Farmer Paul's paddock construction agency will build an animal
  pen. The pen will be in the shape of a rectangle, and gaps that are
  two meters wide will be left open on opposing ends of the
  rectangle. An additional fence will be placed in the middle of the
  rectangle to divide the pen into two equal parts, and the two gaps
  will be opposite to the divider fence. A total of 110 meters of
  fencing will be used. Determine the total width of the pen as a
  function of its height.
\item The interface for a device will consist of two squares.  There
  will be a fancy border around the outside of the two squares. The
  cost of the border will be twenty-five cents per centimetre, and the
  total budget for the borders will be five dollars. Determine the
  length of one side of the first square as a function of the length
  of the second square.
\end{enumerate}
